---
title: "Learning to write functions in R"
author: "Camila Vargas"
date: "5/18/2020"
output: pdf_document
---

This code is a practice code bease on the Data Carpentry lecture on Writing Functions in R.

https://swcarpentry.github.io/r-novice-inflammation/02-func-R/

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


In the last lesson, we learned to combine elements into a vector using the c function, e.g. `x <- c("A", "B", "C")` creates a vector x with three elements. Furthermore, we can extend that vector again using c, e.g. `y <- c(x, "D")` creates a vector y with four elements. Write a function called highlight that takes two vectors as arguments, called content and wrapper, and returns a new vector that has the wrapper vector at the beginning and end of the content:

```{r}

highlight <- function(content, wrapper){
  
  new_vector <- c(wrapper, content, wrapper)
  
  return(new_vector)
  
}

highlight(content = "hello world", wrapper = "***")

```


If the variable v refers to a vector, then v[1] is the vector’s first element and v[length(v)] is its last (the function length returns the number of elements in a vector). Write a function called edges that returns a vector made up of just the first and last elements of its input:

```{r}

edges <- function(input){
  
  output <- c(input[1], input[length(input)])

  return(output)
  }
  
input <-  c("hola", "que", "tal", "po")

edges(input)


```




