---
title: "IntroToTidyverse"
author: "Camila Vargas"
date: "May 15, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```


## Usefull function and how do they work

###gather()

gather(column1, column2, column3, key= "how do I want to call this new", value= "how do I want to call the values from the columns I'm gathering")

###filter()
 If you use the | as and "and"
 != does not equal to.
 
###arrange()
Default arranges the data alphabetically or acending

For opposite order use: desc() or -

###mutate()
  case_when() --> whenever you see an x variable you add y in a new column 
  
  mutate(
  fav_pizza = case_when(
  name== "Sam" ~ "BC",
  name == "An" ~ "P"))

#Practice with data  
```{r}
#Set up
library(here)
library(janitor)

#Read in data
pheno_data <- read_csv("https://raw.githubusercontent.com/samanthacsik/Intro-to-the-Tidyverse/master/data/phenometrics_data.csv")


```
  
### Wrangle:

This dataset is *huge*--we'll want to wrangle it so that it only includes the information that we're interested in. We will:

a. convert variable names to snake_case
b. filter for California observations  
c. select relevant columns of data 
d. rename columns  
e. unite multiple columns  
f. remove any NA values  
g. set the levels for a character vector  

To demonstrate these individual steps, we'll perform each function separately. Notice that we perform subsequent function calls on the data frame generated from the prior step. At the end, we'll show you how to combine all steps into a single, succint code chunk.

 
```{r, include=FALSE}
cali_pheno_simple <- pheno_data %>% 
  clean_names() %>% #converts column names in snake names, this means lowecase and undercore
  filter(state == "CA") %>% 
  select(5:9, phenophase_description, year, month) %>%  ##selecting columns from 5 to 9
  rename(pheno = phenophase_description) %>% 
  unite(genus_species, genus, species, sep = "_") %>% #combines two columns sep= specify what symbol to use to separate the two words.
  drop_na(year, month) %>%  #removes NA
  mutate(pheno = fct_relevel(pheno, c("Flowers or flower buds", "Open flowers", "Pollen release (flowers)")),
         month = fct_relevel(month, month.name)) %>%  ##tells r order this datafram as this given order but it does not actually changes you dataframe, it operate as group_by
  

```
 

With this simplified and cleaned data set, we're ready to explore a subset of the desert species we're most interested in. We love **Joshua trees** (*Yucca brevifolia*), **creosote bushes** (*Larrea tridentata*), and **Mojave yucca** (*Yucca schidigera*) and want to know when these plants are blooming. Let's first isolate data for these species by:

a. filtering for only Joshua tree, creosote bush, and Mojave yucca 
b. grouping observations by month, name, and phenophase
c. finding the total counts by month, name, and phenophase

```{r}

fav_spp <- cali_pheno_simple %>% 
  filter(common_name %in% c("Joshua tree", "creosote bush", "Mojave yucca")) %>%
  group_by(month, common_name, pheno) %>% 
  tally() # you could also use summarize() here!
```

### Plot:

Now that we have our data tallied and in tidy format, we're ready to make a plot! We want to:

a. create a column graph showing the total counts of plants by phenophase and by month
b. create a different panel for each plant species
c. make it pretty

```{r, fig.align = 'center', fig.width = 15, fig.height = 10}
fav_plants_plot <- ggplot(fav_spp, aes(x = month, y = n, fill = pheno)) + # fill = counts of each phenophase
  geom_col(position = "dodge") + # separate columns for each phenophase (instead of stacked)
  facet_wrap(~common_name) + # create separate panels for each species
  labs(x = "Month", y = "Counts", fill = "Phenophase") + # change axis labels and legend names 
  scale_x_discrete(limits = c(month.name)) + # include all months on x-axis, even if there's no data
  scale_y_continuous(expand = c(0,0), breaks = seq(0, 20, by = 3)) + # remove space between columns and x-axis; set y-axis tick mark interval
  scale_fill_manual(values = c("darkseagreen3", "cadetblue")) + # change colors
  theme_classic() + 
  theme(panel.border = element_rect(colour = "black", fill = NA, size = 0.7), 
        axis.text.x = element_text(angle = 45, hjust = 0.9)) 
fav_plants_plot
```



